---
title: "Subnational V-Dem in Colombia: Data wrangle"
author: "Patrick McQuestion"
date: "First version: February 4, 2023. This version: `r format(Sys.Date(), '%d %B %Y')`."
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("G:/My Drive/git/snvdem24")
```

# Mapping Pipeline

First, we decide on the level of analysis for our study. In this case,
we observe that Colombia has fairly complete municipal-level data dating
back to at least the 1950s, at least for a relevant amount of
time-varying variables (e.g., population density, economic development,
voting behavior, etc.).

Second, we collect and format data for mapping purposes. This is a
multi-step process that involves grouping responses in some cases (e.g.,
responses 6-9 that ask to identify 4 cardinal directions). Grouping is
necessary for some variables, but not all. Once grouped, the response
categories themselves must be "translated" into locally relevant
information, which requires a high degree of subjective decision-making.
For example, how many votes or proportion of votes are necessary to
measure "strong" support for the ruling party (question 15)? Or, which
municipalities are considered part of the "North" (question 6) versus
the "West" (question 8)? We privilege the use of granular data that will
allow researchers flexible interpretations of these questions, as well
as subsequent validation from country experts.

Third, we export the response grouping dataframes to .csv format in
order to integrate the datasets into ArcMap software. We do this in
ArcMap, adding the data and joining it to the attribute table for the
base layer (in this case, the shapefile provided by DANE). We then
create and format layers based on the response groupings; for each
response group, we visualize the locations where free and fair elections
are held (and not held), and where civil liberties are stronger (or
weaker).

Fourth, we create a composite image of the layers.

# Data Wrangling

We undertake a series of steps to wrangle data from V-Dem and external
sources. In order to facilitate merging tables to the shapefile maps of
Colombia in ArcMap, we use a key ID (either the official DANE code for
each municipality, or the ArcMap FID number). In the case we prefer to
use the `sf` package, we can proceed to merge along the DANE code.

[Note]{.underline}: This code below concentrates on cross-sectional
data, but it will expand to time-series cross-sectional (\~1900-2022).

```{r, Vdem0, include=FALSE, echo=FALSE}
# make a clean slate
rm(list = ls())
# load needed packages
library(tidyverse)
library(vdemdata)

# Create a data frame of version 13 dataset:
v13 <- vdem
# this df appears to be the same as MC's "full + others" data set: read_dta("C:/Users/mcoppedg/Dropbox/MC/VDemFiles/Archive/V13/V-Dem-CY-Full+Others/V-Dem-CY-Full+Others-v13.dta")

# Filter to just Colombia since 1899:
v13_col <- filter(v13, country_name=="Colombia" & year > 1899)

# Select just the ID and subnational variables
# (Note: v2elsnless and v2elsnmore are not in this df. We will make sure to 
# include them in our request to the data manager.)
v13_col_sn <- v13_col %>%
  select(country_id, country_name, country_text_id, historical_date, year, 
         v2elsnlfc_0:v2elsnlfc_21, v2elsnmrfc_0:v2elsnmrfc_21,
         v2clrgstch_0:v2clrgstch_21, v2clrgwkch_0:v2clrgwkch_21)

# Listing the variable names to check our work:
#names(v13_col_sn)
# Export the result to a CSV file:
#write.csv(v13_col_sn, file = "v13_col_sn.csv")
```

# Exercise: 2018 cross-section

This exercise is designed test the viability of mapping cross-sectional
country-level data before expanding the dataset into panel format.
Therefore, internal data from V-Dem is subset to 2018, and last five
subnational responses are removed because they are not relevant for
Colombia.

```{r, v13_col_2018, include=FALSE, echo=FALSE}
# create subset of V-Dem data for 2018 
v13_col_sn_2018 <- subset(v13_col_sn, year == 2018) 
v13_col_sn_2018 <- v13_col_sn_2018 %>%
  select(-matches("_17$|_18$|_19$|_20$|_21$")) # remove responses not relevant for Colombia
df2rm = c("v13", "v13_col", "v13_col_sn", "df2rm")
rm(list = df2rm)
```

## Table

```{r, v13_col_2018-table, include=FALSE, echo=FALSE}
# Load necessary libraries
library(dplyr)
library(tidyr)
library(knitr)
library(purrr)
library(kableExtra)
library(insight)

selected_vars <- grep("v2elsnlfc|v2elsnmrfc|v2clrgstch|v2clrgwkch", colnames(v13_col_sn_2018), value = TRUE)
selected_df <- v13_col_sn_2018 %>%
  select(all_of(selected_vars))

summary_df <- selected_df %>%
  pivot_longer(cols = everything()) %>%
  mutate(row_id = gsub(".*_(\\d+)$", "\\1", name)) %>%
  mutate(column_id = gsub(".*(v2elsnlfc|v2elsnmrfc|v2clrgstch|v2clrgwkch).*", "\\1", name)) %>%
  pivot_wider(names_from = column_id, values_from = value) %>%
  select(-name) %>%
  mutate(row_id = as.numeric(row_id)) %>% 
  group_by(row_id) %>%
  summarise(across(everything(), ~ifelse(all(is.na(.)), NA, first(na.omit(.))))) %>%
  ungroup()

new_col_names <- c("Elections less free", "Elections more free", "Civil liberties stronger", "Civil liberties weaker")
new_row_names <- c(
  "Rural", "Urban", "Less econ devt", "More econ devt", "Inside capital", 
  "Outside capital", "North", "South", "West", "East", "Civil unrest", 
  "Illicit activity", "Sparse pop.", "Remote", "Indigenous", 
  "Ruling party strong", "Ruling party weak"
)

# Rename columns and modify row_id variable
summary_table <- summary_df %>%
  rename(
    `Elections less free` = v2elsnlfc,
    `Elections more free` = v2elsnmrfc,
    `Civil liberties stronger` = v2clrgstch,
    `Civil liberties weaker` = v2clrgwkch
  ) %>%
  mutate(row_id = new_row_names[as.numeric(row_id) + 1])
summary_table
export_table(summary_table, format = "md")


```

External data from Colombia is drawn from a variety of sources. We merge
this data at the municipal level using DANE codes. These codes are also
used in .dbf tables, such as Colombia's 2018 Census data from DANE. The
Census data Shapefile is useful for initial mapping visualization,
containing not only the administrative boundaries but also contextual
information. The
[dictionary](https://github.com/pmcquest/snvdem24/tree/main/data/geospatial/MGN_ANM_MPIOS)
is in the repository.

```{r, sf1, include=FALSE, echo=FALSE}
library(sf)
col <- st_read("G:/My Drive/git/snvdem24/data/geospatial/2018pmq/BaseLayer/MGN_ANM_MPIOS.shp")
col <- col %>%
  select(1:8)

```

[Note]{.underline}: To facilitate integration of the datasets across
software, we export the Census 2018 attribute table from ArcMaps to .csv
format, then load this data into R, preserving only relevant variables
for merging purposes. While we use DANE codes to merge external data
using R, we use the 2018 Census variable "FID" to merge .csv files into
ArcMap. This is because ArcMaps automatically translates some string
variables (such as the DANE codes) into a numeric ones. For this reason,
we integrate the FID variable for the individual .csv files we write.

```{r, Arc1, include=FALSE, echo=FALSE}
library(readxl)
MGN18 <- read_excel("G:/My Drive/git/snvdem24/data/geospatial/MGN_ANM_MPIOS/MGN18.xls") 
# We select only relevant variables for merging purposes.
# colnames(MGN18)
MGN18 <- MGN18 %>%
  select(1:7)

# we merge the FID variable to our main shapefile
col <- merge(col, MGN18[, c("MPIO_CDPMP", "FID")], by = "MPIO_CDPMP", all.x = TRUE)
```

## 0-1: Rurality

The survey asks about elections and civil liberties in rural or urban
locations. Rurality or urbanization can be measured in many ways. For
example, the World Bank's [Rural Access Index
(RAI)](https://datacatalog.worldbank.org/search/dataset/0038250) or
European Commission's [Global Human Settlement Layer
(GHSL)](https://ghsl.jrc.ec.europa.eu/CFS.php) are exemplary. Similarly,
[Waldorf and Kim (2015)](https://purr.purdue.edu/publications/2960/1)
create an index that considers population size, density, remoteness, and
built-up areas. [DANE
(2015)](https://colaboracion.dnp.gov.co/CDT/Estudios%20Econmicos/2015ago6%20Documento%20de%20Ruralidad%20-%20DDRS-MTC.pdf)
uses a classification based on OECD criteria that defines rural
territories as those with 150 or fewer persons per kilometer squared,
finding that over 75% of Colombian municipalities do not meet the
threshold of 100 persons per km\^2. From another perspective, DANE
estimates that close to 4/5 of the population lives in urban centers. In
this exercise, we take a demographic approach that considers the
proportion of the municipal rural population over the total municipal
population. This basic "Rurality Index" measure created by CEDE uses
Census data (DANE).

```{r, V18_0-1_read, include=FALSE, echo=FALSE}
# municipal-level data for distance to capital city
IR18 <- read_excel("G:/My Drive/git/snvdem24/data/geospatial/2018pmq/0-1_Rurality/PANEL_CARACTERISTICAS_GENERALES(2021).xlsx")
colnames(IR18)[3] = "MPIO_CDPMP"
colnames(IR18)[7] = "Year"
IR18 <- IR18 %>%
  filter(Year == 2018)
# this data set also contains important descriptive information (department, province) that we will hang onto
IR18 <- IR18[c("MPIO_CDPMP", "depto", "provincia", "indrural")]

# Because the import creates a numeric field for DANE code, we must convert this numeric variable to character and then assure each observation has the corresponding 5 digits
IR18$MPIO_CDPMP <- as.character(IR18$MPIO_CDPMP)
# Add a 0 before values with 4 digits only
IR18$MPIO_CDPMP <- ifelse(nchar(IR18$MPIO_CDPMP) == 4, paste0("0", IR18$MPIO_CDPMP), IR18$MPIO_CDPMP)

```

The survey asks two questions, however, the systematized concept can
take many forms. In this case, our V-Dem scores reflect the proportion
of respondents who selected rural, urban, or both categories to describe
election fairness and civil liberties. We can interact these proportions
per question with the rurality index from CEDE (also a proportional
score) to provide a more granular view of democracy at the municipal
level based on our particular systematization. Another option is to
create a dichotomous threshold between urban and rural variables and
assign V-Dem scores accordingly. We can take the mid-point as a
threshold for rural and urban municipalities: if more than half of the
population is considered rural inhabitants, then the municipality is
considered rural. This can be adjusted later if necessary. - As a first
cut a good way to start

```{r, V18_0-1_vdem-scale, include=FALSE, echo=FALSE}
# Option #1: interact V-Dem with rural index score
#For example, a high rural index multiplied by a high proportion of V-Dem's response will produce correspondingly high scores. This may occur with the values for less free and fair elections for rural (.875) vs. urban (.125) areas.
# ([indrur0t1]*Rural)+((1-[indrur0t1])*Urban)
# .80*.875 + (1-.8)*.125
IR18 <- IR18 %>% 
  mutate(el_0t1x = (indrural*v13_col_sn_2018$v2elsnlfc_0)+((1-indrural)*v13_col_sn_2018$v2elsnlfc_1)) %>%
  mutate(em_0t1x = (indrural*v13_col_sn_2018$v2elsnmrfc_0)+((1-indrural)*v13_col_sn_2018$v2elsnmrfc_1)) %>%
  mutate(cs_0t1x = (indrural*v13_col_sn_2018$v2clrgstch_0)+((1-indrural)*v13_col_sn_2018$v2clrgstch_1)) %>%
  mutate(cw_0t1x = (indrural*v13_col_sn_2018$v2clrgwkch_0)+((1-indrural)*v13_col_sn_2018$v2clrgwkch_1))

# Min-max normalization function
nz <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}
# Create new variables with normalized values. Normalization may be important when comparing across variable scales.
df_nz <- data.frame(el_0t1xs = nz(IR18$el_0t1x),
                    em_0t1xs = nz(IR18$em_0t1x),
                    cs_0t1xs = nz(IR18$cs_0t1x),
                    cw_0t1xs = nz(IR18$cw_0t1x))
# Combine the normalized variables with the original dataframe
IR18 <- cbind(IR18, df_nz)

# For sf (R), merge this data to shapefile
col <- merge(col, IR18, by = "MPIO_CDPMP", all.x = TRUE)


```

```{r, V18_0-1_vdem-dicho, include=FALSE, echo=FALSE}
# Option #2: we dichotomize the rurality score and assign the corresponding V-Dem score
# "less free and fair subnational elections"
IR18 <- IR18 %>% 
  mutate(elsnl_0t1 = case_when(
    indrural > 0.50 ~ v13_col_sn_2018$v2elsnlfc_0,  
    indrural < 0.50 ~ v13_col_sn_2018$v2elsnlfc_1,
    TRUE ~ NA_real_
))
# "more free and fair subnational elections"
IR18 <- IR18 %>% 
  mutate(elsnmr_0t1 = case_when(
    indrural > 0.50 ~ v13_col_sn_2018$v2elsnmrfc_0,
    indrural < 0.50 ~ v13_col_sn_2018$v2elsnmrfc_1,
    TRUE ~ NA_real_ 
))
# "stronger civil liberties"
IR18 <- IR18 %>% 
  mutate(clrgst_0t1 = case_when(
    indrural > 0.50 ~ v13_col_sn_2018$v2clrgstch_0, 
    indrural < 0.50 ~ v13_col_sn_2018$v2clrgstch_1,
    TRUE ~ NA_real_
))
# "weaker civil liberties"
IR18 <- IR18 %>% 
  mutate(clrgwk_0t1 = case_when( 
    indrural > 0.50 ~ v13_col_sn_2018$v2clrgwkch_0,  
    indrural < 0.50 ~ v13_col_sn_2018$v2clrgwkch_1,
TRUE ~ NA_real_
))


```

```{r, V18_0-1_vdem-plot, include=FALSE, echo=FALSE}

ggplot() +
  geom_sf(data = col, aes(fill = el_0t1xs)) + # just to check results
  scale_fill_viridis_c() +  # Adjust color scale as needed
  theme_minimal()

```

```{r, V18_0-1_export, include=FALSE, echo=FALSE}
# For sf (R), merge this data to shapefile
col <- merge(col, IR18, by = "MPIO_CDPMP", all.x = TRUE)


# merge base-layer dataset to obtain FID variable
IR18 <- merge(IR18, MGN18[, c("MPIO_CDPMP", "FID")], by = "MPIO_CDPMP", all.x = TRUE)
IR18 <- IR18 %>%
  rename(indrur0t1 = indrural) %>%
  rename(FID_0t1 = FID)

# For ArcMaps, export the final merge to a .csv file
write.csv(IR18, file = "G:/My Drive/git/snvdem24/data/geospatial/2018pmq/0-1_Rurality/IR18_0-1.csv", row.names = FALSE)

#clean things up for next variable
df2rm = c("IR18", "df2rm")
rm(list = df2rm)

```

## 2-3: Economic development

[DANE](https://www.dane.gov.co/index.php/estadisticas-por-tema/cuentas-nacionales/cuentas-nacionales-departamentales)
provides data for Gross Domestic Product (GDP), known as Producto Bruto
Interno (PBI) per capita. They offer a historical "retropolation" of
department-level PBI dating back to 1980, and a municipal-level measure
of "value added" dating back to 2011 (methodological documentation is
included in the repository). Values are in Colombian pesos (COP) and
scaled to \$1,000,000,000.

The metropolition data extends farther back in time, making it more
amenable to panel data analysis. Also, there may also be reduced
measurement error given the difficulty of isolating economic
productivity to municipalities.

Most departments are covered by the retropolation measure, but newer
departments (many of which were inducted in 1991) are grouped into one
unit, and therefore cannot be mapped based on DANE's Departmental Code
system without data intervention. These are: Amazonas (91), Arauca (81),
Casanare (85), Guainía (94), Guaviare (95), Putumayo (86), San Andrés,
Providencia y Santa Catalina (Archipiélago) (88), Vaupés (97) y Vichada
(99).

Nearly all municipalities are measured by the Municipal Value Added
(VAM) measure, however, data only exists from 2011 onward.

For experimental purposes, we try both measures for 2018.

```{r, V18_2-3_read, include=FALSE, echo=FALSE}
# municipal-level data for "value added"
VAM18 <- read_excel("G:/My Drive/git/snvdem24/data/geospatial/2018pmq/2-3_EconDevt/anexo-2020-2021-provisional-valor-agregado-municipio-2011-2021.xlsx", sheet = "Cuadro 9", range = "A11:I1133")
# rename columns to facilitate merging shapefile
VAM18 <- VAM18 %>%
  rename(MPIO_CDPMP = `Código Municipio`) %>%
  rename(DPTO_CCDGO = `Código Departamento`) %>%
  rename(VAM18_2t3 = `Valor agregado\r\n`) %>%
  select(1|3|8)

# merge base-layer dataset to obtain FID variable (for later integration with ArcMaps)
ED18 <- merge(VAM18, MGN18[, c("MPIO_CDPMP", "FID")], by = "MPIO_CDPMP", all.x = TRUE)
ED18 <- ED18 %>%
  rename(FID_2t3 = FID)

# department-level data for "PBI per capita"
PBID <- read_excel("G:/My Drive/git/snvdem24/data/geospatial/2018pmq/2-3_EconDevt/anex-PIBDep-RetropolacionDepartamento-2022pr.xlsx", sheet = "Cuadro 1", range = "A10:AS36")
PBID <- PBID %>%
  rename(DPTO_CCDGO = `Código Departamento (DIVIPOLA)`) %>%
  select(DPTO_CCDGO, `2018`) %>% #keep only 2018 data 
  rename(PBID18_2t3 = `2018`)

ED18 <- merge(ED18, PBID, by = "DPTO_CCDGO", all.x = TRUE)

```

After creating the data frame for economic development at municipal and
departmental levels, we must conceptualize "less" versus "more"
developed areas before merging with V-Dem data. A simple option could be
to dichotomize economic development based on median values. A more
complex method would interact these municipal measures with V-Dem
responses, as we did with rurality, but this would involve scaling and
other steps beforehand.

```{r, V18_2-3_vdem_median, include=FALSE, echo=FALSE}
median(ED18$VAM18_2t3) #municipal threshold as median: 137.6316
median(ED18$PBID18_2t3, na.rm = TRUE) #departmental threshold as median: 26884.21 (with NA values excluded)
## We may want to use quintile power, but we don't want to assign too much weight to these categories. 

#Q2: in "areas that are less economically developed" there are ...
## at the municipal level...
ED18 <- ED18 %>% 
  #"less free and fair subnational elections" 
  mutate(elsnl_2m  = case_when( #Reduced the variable names to less than 10 characters in order to avoid merging issues in ArcMap
    VAM18_2t3< 138 ~ v13_col_sn_2018$v2elsnlfc_2, TRUE ~ NA_real_)) %>%
  #"more free and fair subnational elections"
  mutate(elsnmr_2m  = case_when(
    VAM18_2t3< 138 ~ v13_col_sn_2018$v2elsnmrfc_2, TRUE ~ NA_real_)) %>%
  #"stronger civil liberties" 
  mutate(clrgst_2m  = case_when(
    VAM18_2t3< 138 ~ v13_col_sn_2018$v2clrgstch_2, TRUE ~ NA_real_)) %>%
  #"weaker civil liberties" 
  mutate(clrgwk_2m  = case_when(
    VAM18_2t3< 138 ~ v13_col_sn_2018$v2clrgwkch_2, TRUE ~ NA_real_)) %>% 
## at the departmental level
  #"less free and fair subnational elections" 
  mutate(elsnl_2d = case_when(
    PBID18_2t3< 26884 ~ v13_col_sn_2018$v2elsnlfc_2, TRUE ~ NA_real_)) %>%
  #"more free and fair subnational elections"
  mutate(elsnmr_2d = case_when(
    PBID18_2t3< 26884 ~ v13_col_sn_2018$v2elsnmrfc_2, TRUE ~ NA_real_)) %>%
  #"stronger civil liberties" 
  mutate(clrgst_2d = case_when(
    PBID18_2t3< 26884 ~ v13_col_sn_2018$v2clrgstch_2, TRUE ~ NA_real_)) %>%
  #"weaker civil liberties" 
  mutate(clrgwk_2d = case_when(
    PBID18_2t3< 26884 ~ v13_col_sn_2018$v2clrgwkch_2, TRUE ~ NA_real_))

#Q3: in "areas that are more economically developed" there are ...
## at the municipal level...
ED18 <- ED18 %>% 
  #"less free and fair subnational elections" 
  mutate(elsnl_3m  = case_when(
    VAM18_2t3> 138 ~ v13_col_sn_2018$v2elsnlfc_3, TRUE ~ NA_real_)) %>%
  #"more free and fair subnational elections"
  mutate(elsnmr_3m  = case_when(
    VAM18_2t3> 138 ~ v13_col_sn_2018$v2elsnmrfc_3, TRUE ~ NA_real_)) %>%
  #"stronger civil liberties" 
  mutate(clrgst_3m  = case_when(
    VAM18_2t3> 138 ~ v13_col_sn_2018$v2clrgstch_3, TRUE ~ NA_real_)) %>%
  #"weaker civil liberties" 
  mutate(clrgwk_3m  = case_when(
    VAM18_2t3> 138 ~ v13_col_sn_2018$v2clrgwkch_3, TRUE ~ NA_real_)) %>% 
## at the departmental level
  #"less free and fair subnational elections" 
  mutate(elsnl_3d = case_when(
    PBID18_2t3> 26884 ~ v13_col_sn_2018$v2elsnlfc_3, TRUE ~ NA_real_)) %>%
  #"more free and fair subnational elections"
  mutate(elsnmr_3d = case_when(
    PBID18_2t3> 26884 ~ v13_col_sn_2018$v2elsnmrfc_3, TRUE ~ NA_real_)) %>%
  #"stronger civil liberties" 
  mutate(clrgst_3d = case_when(
    PBID18_2t3> 26884 ~ v13_col_sn_2018$v2clrgstch_3, TRUE ~ NA_real_)) %>%
  #"weaker civil liberties" 
  mutate(clrgwk_3d = case_when(
    PBID18_2t3> 26884 ~ v13_col_sn_2018$v2clrgwkch_3, TRUE ~ NA_real_))
```

Unclear how to set up the interaction option.

```{r, V18_2-3_vdem_x, include=FALSE, echo=FALSE}
#municipal estimates: 
ED18 <- ED18 %>% 
  mutate(el_2t3xm = (VAM18_2t3*v13_col_sn_2018$v2elsnlfc_2)+((1-VAM18_2t3)*v13_col_sn_2018$v2elsnlfc_3)) %>%
  mutate(em_2t3xm = (VAM18_2t3*v13_col_sn_2018$v2elsnmrfc_2)+((1-VAM18_2t3)*v13_col_sn_2018$v2elsnmrfc_3)) %>%
  mutate(cs_2t3xm = (VAM18_2t3*v13_col_sn_2018$v2clrgstch_2)+((1-VAM18_2t3)*v13_col_sn_2018$v2clrgstch_3)) %>%
  mutate(cw_2t3xm = (VAM18_2t3*v13_col_sn_2018$v2clrgwkch_2)+((1-VAM18_2t3)*v13_col_sn_2018$v2clrgwkch_3))

#departmental estimates
ED18 <- ED18 %>% 
  mutate(el_2t3xd = (PBID18_2t3*v13_col_sn_2018$v2elsnlfc_2)+((1-PBID18_2t3)*v13_col_sn_2018$v2elsnlfc_3)) %>%
  mutate(em_2t3xd = (PBID18_2t3*v13_col_sn_2018$v2elsnmrfc_2)+((1-PBID18_2t3)*v13_col_sn_2018$v2elsnmrfc_3)) %>%
  mutate(cs_2t3xd = (PBID18_2t3*v13_col_sn_2018$v2clrgstch_2)+((1-PBID18_2t3)*v13_col_sn_2018$v2clrgstch_3)) %>%
  mutate(cw_2t3xd = (PBID18_2t3*v13_col_sn_2018$v2clrgwkch_2)+((1-PBID18_2t3)*v13_col_sn_2018$v2clrgwkch_3))

```

```{r, V18_2-3_export, include=FALSE, echo=FALSE}
# For sf (R), merge this data to shapefile
# but first drop the DPTO variable
ED18_2t3 <- subset(ED18_2t3, select = -DPTO_CCDGO)
col <- merge(col, ED18_2t3, by = "MPIO_CDPMP", all.x = TRUE)

# For ArcMap, export V-Dem merged data to a .csv file that will be joined to map
write.csv(ED18_2t3, file = "G:/My Drive/git/snvdem24/data/geospatial/2018pmq/2-3_EconDevt/ED18_2-3.csv", row.names = FALSE)

#clean things up for next variable
df2rm = c("ED18", "ED18_2t3", "PBID", "VAM18", "df2rm")
rm(list = df2rm)
```

## 4-5: Inside or outside of capital city

The survey asks about elections and civil liberties "Inside / Outside
the capital city". In the context of V-Dem this likely refers to the
national capital (Bogota), but it could also refer to municipal
districts that are also department capitals. CEDE contains 2 variables
of interest for systematization of this concept: distance from Bogota
and distance from department capital. Even though we can assume
respondents were considering the national capital, We will keep both in
case we consider the alternative interpretation. Note: It may be
possible to calculate this variable in ArcMaps using the field
calculator, but this requires advanced techniques.

```{r, V18_4-5_read, include=FALSE, echo=FALSE}
# municipal-level data for distance to capital city
CEDE18 <- read_excel("G:/My Drive/git/snvdem24/data/geospatial/2018pmq/4-5_DisCapital/PANEL_CARACTERISTICAS_GENERALES(2021).xlsx")
colnames(CEDE18)[3] = "MPIO_CDPMP"
colnames(CEDE18)[7] = "Year"
CEDE18 <- CEDE18 %>%
  filter(Year == 2018)
# this data set also contains important descriptive information (department, province, rurality index) that we will hang onto
CEDE18 <- CEDE18[c("MPIO_CDPMP", "discapital", "disbogota")]

# Because the import creates a numeric field for DANE code, we must convert this numeric variable to character and then assure each observation has the corresponding 5 digits
CEDE18$MPIO_CDPMP <- as.character(CEDE18$MPIO_CDPMP)
# Add a 0 before values with 4 digits only
CEDE18$MPIO_CDPMP <- ifelse(nchar(CEDE18$MPIO_CDPMP) == 4, paste0("0", CEDE18$MPIO_CDPMP), CEDE18$MPIO_CDPMP)

```

Although the survey asks two questions, technically these responses can
be grouped into one dichotomous variable. We will therefore create one
variable for each survey pair of survey questions that assign
dichotomous values (0=capital, \>0=non-capital). Nonetheless, we can
also consider the relevance of proximity following a continuous scale.
We therefore calculate this scale for the national capital distance
using the range.

```{r, V18_4-5_vdem-dicho, include=FALSE, echo=FALSE}
# National capital:
# "less free and fair subnational elections"
CEDE18 <- CEDE18 %>% 
  mutate(el_4t5n = case_when(
    disbogota == 0 ~ v13_col_sn_2018$v2elsnlfc_4,  
    disbogota > 0 ~ v13_col_sn_2018$v2elsnlfc_5,
    TRUE ~ NA_real_
))
# "more free and fair subnational elections"
CEDE18 <- CEDE18 %>% 
  mutate(em_4t5n = case_when(
    disbogota == 0 ~ v13_col_sn_2018$v2elsnmrfc_4,
    disbogota > 0 ~ v13_col_sn_2018$v2elsnmrfc_5,
    TRUE ~ NA_real_ 
))
# "stronger civil liberties"
CEDE18 <- CEDE18 %>% 
  mutate(cs_4t5n = case_when(
    disbogota == 0 ~ v13_col_sn_2018$v2clrgstch_4, 
    disbogota > 0 ~ v13_col_sn_2018$v2clrgstch_5,
    TRUE ~ NA_real_
))
# "weaker civil liberties"
CEDE18 <- CEDE18 %>% 
  mutate(cw_4t5n = case_when( 
    disbogota == 0 ~ v13_col_sn_2018$v2clrgwkch_4,  
    disbogota > 0 ~ v13_col_sn_2018$v2clrgwkch_5,
TRUE ~ NA_real_
))

# Department capital:
# "less free and fair subnational elections"
CEDE18 <- CEDE18 %>% 
  mutate(el_4t5d = case_when(
    discapital == 0 ~ v13_col_sn_2018$v2elsnlfc_4,  
    discapital > 0 ~ v13_col_sn_2018$v2elsnlfc_5,
    TRUE ~ NA_real_
))
# "more free and fair subnational elections"
CEDE18 <- CEDE18 %>% 
  mutate(em_4t5d = case_when(
    discapital == 0 ~ v13_col_sn_2018$v2elsnmrfc_4,
    discapital > 0 ~ v13_col_sn_2018$v2elsnmrfc_5,
    TRUE ~ NA_real_ 
))
# "stronger civil liberties"
CEDE18 <- CEDE18 %>% 
  mutate(cs_4t5d = case_when(
    discapital == 0 ~ v13_col_sn_2018$v2clrgstch_4, 
    discapital > 0 ~ v13_col_sn_2018$v2clrgstch_5,
    TRUE ~ NA_real_
))
# "weaker civil liberties"
CEDE18 <- CEDE18 %>% 
  mutate(cw_4t5d = case_when( 
    discapital == 0 ~ v13_col_sn_2018$v2clrgwkch_4,  
    discapital > 0 ~ v13_col_sn_2018$v2clrgwkch_5,
TRUE ~ NA_real_
))
```

```{r, V18_4-5_vdem-scale, include=FALSE, echo=FALSE}

# continuous scale
range(CEDE18$disbogota) #1270.85 is furthest distance
CEDE18$pct_4t5x <- CEDE18$disbogota / 1270.85 #create a proportional score of distance from Bogota 

CEDE18 <- CEDE18 %>% 
  mutate(el_4t5x = ((1-pct_4t5x)*v13_col_sn_2018$v2elsnlfc_4)+(pct_4t5x*v13_col_sn_2018$v2elsnlfc_5)) %>%
  mutate(em_4t5x = ((1-pct_4t5x)*v13_col_sn_2018$v2elsnmrfc_4)+(pct_4t5x*v13_col_sn_2018$v2elsnmrfc_5)) %>%
  mutate(cs_4t5x = ((1-pct_4t5x)*v13_col_sn_2018$v2clrgstch_4)+(pct_4t5x*v13_col_sn_2018$v2clrgstch_5)) %>%
  mutate(cw_4t5x = ((1-pct_4t5x)*v13_col_sn_2018$v2clrgwkch_4)+(pct_4t5x*v13_col_sn_2018$v2clrgwkch_5))

# Min-max normalization function
nz <- function(x) {
  return((x - min(x)) / (max(x) - min(x)))
}
# Create new variables with normalized values. Normalization may be important when comparing across variable scales.
df_nz <- data.frame(el_4t5xs = nz(CEDE18$el_4t5x),
                    em_4t5xs = nz(CEDE18$em_4t5x),
                    cs_4t5xs = nz(CEDE18$cs_4t5x),
                    cw_4t5xs = nz(CEDE18$cw_4t5x))
# Combine the normalized variables with the original dataframe
CEDE18 <- cbind(CEDE18, df_nz)

# For sf (R), merge this data to shapefile
col <- merge(col, CEDE18, by = "MPIO_CDPMP", all.x = TRUE)
```

```{r, V18_4-5_vdem-plot, include=FALSE, echo=FALSE}

ggplot() +
  geom_sf(data = col, aes(fill = el_4t5x)) + # just to check results
  scale_fill_viridis_c() +  # Adjust color scale as needed
  theme_minimal()

```

```{r, V18_4-5_export, include=FALSE, echo=FALSE}
# For sf (R), merge this data to shapefile
col <- merge(col, CEDE18, by = "MPIO_CDPMP", all.x = TRUE)


# merge base-layer dataset to obtain FID variable
CEDE18 <- merge(CEDE18, MGN18[, c("MPIO_CDPMP", "FID")], by = "MPIO_CDPMP", all.x = TRUE)
CEDE18 <- CEDE18 %>%
  rename(discap4t5 = discapital) %>%
  rename(disbog4t5 = disbogota) %>% 
  rename(FID_4t5 = FID)

# For ArcMaps, export the final merge to a .csv file
write.csv(CEDE18, file = "G:/My Drive/git/snvdem24/data/geospatial/2018pmq/4-5_DisCapital/DC18_4-5.csv", row.names = FALSE)

#clean things up for next variable
df2rm = c("CEDE18", "df2rm")
rm(list = df2rm)

```

## 6-9: Cardinal directions

In order to map the four cardinal regions in Colombia, we are guided by
external data from DANE who recently classified the country into 6
[macro-regions](https://www.dane.gov.co/index.php/estadisticas-por-tema/informacion-regional/informacion-estadistica-desagregada-con-enfoque-territorial-y-diferencial/informacion-del-dane-para-la-toma-de-decisiones-en-departamentos-y-ciudades-capitales).
We find a public-access table of municipalities classified into these
regions
[here](https://www.datos.gov.co/Mapas-Nacionales/Departamentos-y-municipios-de-Colombia/xdk5-pm3f/about_data).
There are other divisions, such as the 5 regions used by CEDE studies
(Andina, Caribe, Pacifica, Orinoquia, Amazonia), but for now we go with
the prior classification.

```{r, V18_6-9_read, include=FALSE, echo=FALSE}
#data from DANE regions requires some formatting before merging
reg <- read_csv("G:/My Drive/git/snvdem24/data/geospatial/2018pmq/6-9_Cardinal/Departamentos_y_municipios_de_Colombia_20240312.csv")
reg <- reg %>%
  rename(mpio = `CÓDIGO DANE DEL MUNICIPIO`)
reg$MPIO_CDPMP <- sprintf("%.3f", reg$mpio) #Codigo municipio should have 5 digits
reg$MPIO_CDPMP <- gsub("\\.", "", reg$MPIO_CDPMP)
reg$MPIO_CDPMP <- ifelse(nchar(reg$MPIO_CDPMP) == 4, sprintf("0%s", reg$MPIO_CDPMP), reg$MPIO_CDPMP)

```

We decide to group the regions in the following manner (see code below),
creating a new variable called "Cardinal". Then, we merge our V-Dem data
related to responses #6-9 for each subnational survey question.

```{r, V18_6-9_vdem, include=FALSE, echo=FALSE}
unique(reg$REGION)
reg <- reg %>%
  mutate(Cardinal = case_when(
    REGION %in% c('Región Caribe', 'Región Eje Cafetero - Antioquia') ~ 'North',
    REGION %in% c('Región Centro Oriente', 'Región Llano') ~ 'East',
    REGION == 'Región Centro Sur' ~ 'South',
    REGION == 'Región Pacífico' ~ 'West',
    TRUE ~ NA_character_
  ))

# V-Dem merge:
# "less free and fair subnational elections"
reg <- reg %>% 
  mutate(elsnlfc = case_when(
    Cardinal == "North" ~ v13_col_sn_2018$v2elsnlfc_6,  
    Cardinal == "South" ~ v13_col_sn_2018$v2elsnlfc_7,
    Cardinal == "West" ~ v13_col_sn_2018$v2elsnlfc_8,  
    Cardinal == "East" ~ v13_col_sn_2018$v2elsnlfc_9,
    TRUE ~ NA_real_ 
))
# "more free and fair subnational elections"
reg <- reg %>% 
  mutate(elsnmrfc = case_when(
    Cardinal == "North" ~ v13_col_sn_2018$v2elsnmrfc_6,
    Cardinal == "South" ~ v13_col_sn_2018$v2elsnmrfc_7,
    Cardinal == "West" ~ v13_col_sn_2018$v2elsnmrfc_8,  
    Cardinal == "East" ~ v13_col_sn_2018$v2elsnmrfc_9,
    TRUE ~ NA_real_ 
))
# "stronger civil liberties"
reg <- reg %>% 
  mutate(clrgstch = case_when(
    Cardinal == "North" ~ v13_col_sn_2018$v2clrgstch_6, 
    Cardinal == "South" ~ v13_col_sn_2018$v2clrgstch_7,
    Cardinal == "West" ~ v13_col_sn_2018$v2clrgstch_8,  
    Cardinal == "East" ~ v13_col_sn_2018$v2clrgstch_9,
    TRUE ~ NA_real_
))
# "weaker civil liberties"
reg <- reg %>% 
  mutate(clrgwkch = case_when( 
    Cardinal == "North" ~ v13_col_sn_2018$v2clrgwkch_6,  
    Cardinal == "South" ~ v13_col_sn_2018$v2clrgwkch_7,
    Cardinal == "West" ~ v13_col_sn_2018$v2clrgwkch_8,  
    Cardinal == "East" ~ v13_col_sn_2018$v2clrgwkch_9,
TRUE ~ NA_real_ 
))

# Select and rename variables
reg <- reg %>% 
  select(1|6:11) %>%
  rename(Card_6t9 = Cardinal) %>%
  rename(elsnl_6t9 = elsnlfc) %>%
  rename(elsnmr_6t9 = elsnmrfc) %>%
  rename(clrgst_6t9 = clrgstch) %>%
  rename(clrgwk_6t9 = clrgwkch)

```

```{r, V18_6-9_export, include=FALSE, echo=FALSE}

# For sf (R), merge this data to shapefile
col <- merge(col, reg, by = "MPIO_CDPMP", all.x = TRUE)

# For ArcMaps, merge base-layer dataset to obtain FID variable (for later integration with ArcMaps)
reg <- merge(reg, MGN18[, c("MPIO_CDPMP", "FID")], by = "MPIO_CDPMP", all.x = TRUE)
reg <- reg %>% 
  rename(FID_6t9 = FID)
#Belen de Bajira (Choco) does not have an FID number, which causes problems in ArcMap when merging. This municipality was only recently inaugurated.
reg <- reg[complete.cases(reg$FID_6t9), ] 

# Export the final merge to a .csv file that can be read into ArcMaps:
write.csv(reg, file = "G:/My Drive/git/snvdem24/data/geospatial/2018pmq/6-9_Cardinal/CD18_6-9.csv", row.names = FALSE)

#clean things up for next variable
df2rm = c("reg", "df2rm")
rm(list = df2rm)

```

## 10: Civil unrest

## 11: Illicit activity

## 12: Sparse population density

-   will likely covary with rurality

## 13: Remoteness

-   will likely covary with rurality

## 14: Indigenous

## 15-16: National ruling party

The questions ask "Areas where national ruling party or group is strong
/ weak." The most straightforward measure is a dichotomous variable
where president receives a majority of the vote or not. A more
variegated measure can leverage the percentage of votes cast for the
president-elect vs. runner-up, for example, however this measure may
need to account for blank votes (which may signify protest votes).

The measure for ruling party support used here will be: % of vote that
supports ruling party. For first wrangle, I will choose 2018 runoff
election (second round) between Petro and Duque (Duque won, so his is
ruling party). This makes things easier because there are only 2
candidates, the winner and runner-up. Elections without second rounds
(prior to 1994, as well as 2002 and 2006) may require additional coding.

```{r, V18_15-16_read, include=FALSE, echo=FALSE}
# Import voting data from Registraduria (RNEC).
library(readr)
rp18 <- read_csv("G:/My Drive/git/snvdem24/data/geospatial/2018pmq/15-16_RulingParty/2018_presidencia_segunda_vuelta_dta_c27d4515ed.csv") # this long dataset contains only 2 candidates because it was a second-round election. other datasets will include more rows depending on the number of candidates running.

#Codigo municipio should have 5 digits
rp18$MPIO_CDPMP <- as.character(rp18$codmpio)
rp18$MPIO_CDPMP <- ifelse(nchar(rp18$MPIO_CDPMP) == 4, sprintf("0%s", rp18$MPIO_CDPMP), rp18$MPIO_CDPMP)
# merge base-layer dataset to obtain FID variable (for later integration with ArcMaps)
rp18 <- merge(rp18, MGN18[, c("MPIO_CDPMP", "FID")], by = "MPIO_CDPMP", all.x = TRUE)
```

The voting data requires more steps for cleaning. In this exercise, we
will take all possible vote entries: top 2 candidates, as well as no
mark, blank, or null ballots. We see that Duque won by high margin (over
26% on average) in 2018.

```{r, V18_15-16_edit, include=FALSE, echo=FALSE}
library(dplyr)
# Step 1: Calculate total votes cast at municipal level (allows for percentages later on).
rp18 <- rp18 %>%
  group_by(FID) %>%
  mutate(votototal = sum(votos)) %>%
  mutate(vtpc = votos/votototal) %>% #optional: useful for quick look at percentages, SD, etc. 
  mutate(vtpc0 = round(vtpc*100, 2)) 

# For mapping at the municipal level, values may take the form of: percent difference in votes between top winner and runner-up candidate(s). This requires a few steps:

#Step 2: Pivot municipal voting behavior into wide format.
p18d0 <- rp18 %>%
  group_by(FID, codigo_lista) %>%
  pivot_wider(names_from = codigo_lista, values_from = votos) %>% # Pivot to wide format
  rename(Petro = `1`,
         Duque = `2`,
         nomark = `997`,
         null = `998`,
         blank = `999`)

## Step 3: Create new dfs for voting behavior at municipal level -- is there a more efficient code?
p18d2 <- p18d0 %>%
  select(FID, Petro) %>%
  filter(!is.na(Petro))
p18d2 <- p18d2[!duplicated(p18d2$FID), ]
p18d3 <- p18d0 %>%
  select(FID, Duque) %>%
  filter(!is.na(Duque))
p18d3 <- p18d3[!duplicated(p18d3$FID), ]
p18d4 <- p18d0 %>%
  select(FID, nomark) %>%
  filter(!is.na(nomark))
p18d4 <- p18d4[!duplicated(p18d4$FID), ]
p18d5 <- p18d0 %>%
  select(FID, null) %>%
  filter(!is.na(null))
p18d5 <- p18d5[!duplicated(p18d5$FID), ]
p18d6 <- p18d0 %>%
  select(FID, blank) %>%
  filter(!is.na(blank))
p18d6 <- p18d6[!duplicated(p18d6$FID), ]
p18d7 <- p18d0 %>%
  group_by(FID) %>%
  summarise(vtotal = first(votototal))
p18d7 <- p18d7[!duplicated(p18d7$FID), ]

RPs <- list(p18d2, p18d3, p18d4, p18d5, p18d6, p18d7)
merged_RP <- Reduce(function(x, y) merge(x, y, by = "FID", all = TRUE), RPs)

df2rm2 = c("rp18", "p18d0", "p18d2", "p18d3", "p18d4", "p18d5", "p18d6", "p18d7", "df2rm2", "RPs")
rm(list = df2rm2)

## Step 3: Create margin of support variable for winner and runner-up candidates. Positive value indicates support for winner; negative indicates opposition. Variable order may vary depending on the election year data from RNEC
RP_15 <- merged_RP %>%
  mutate(MOV_top2 = Duque-Petro) %>% #Raw vote-count margin of victory
  mutate(Duque_pct = Duque/vtotal) %>%
  mutate(Petro_pct = Petro/vtotal) %>%
  mutate(MOV_pct = Duque_pct-Petro_pct) #Standardized percent margin of victory

```

After creating the data frame for voting behavior at the municipal
level, we must decide on the threshold for "strong" versus "weak"
support of the ruling party. We could take one standard deviation or
higher as the threshold, but even so: what is the criteria? In the
interest of simplicity, we will create a dichotomous measure, with
margin of victory (MOV) above 10 percentage points, applied in either
direction (positive for winner Duque; negative for loser Petro). This
seems like a standard difference, but further research could guide this
threshold setting. - ideally, don't dichotomize. Use a continuous
scale. - cl stronger where RP is stronger and experts agree thats

```{r, V18_15-16_vdem, include=FALSE, echo=FALSE}
mean(RP_15$MOV_top2_pct) #Duque margin is 26.3% on average (quite high).
sd(RP_15$MOV_top2_pct) #SD for the margin is larger than the mean (39.7%) suggesting abnormal distribution
hist(RP_15$MOV_pct, main = "Histogram of MOV", xlab = "MOV", col = "skyblue", border = "black") #left-tail long (margin for Petro)


#Q15: in "areas where support [10% MOV] for the national ruling party is strong" there are ...
## instead of arbitrary cut-point
RP_15 <- RP_15 %>% 
  #"less free and fair subnational elections" 
  mutate(elsnl_15 = case_when(
    MOV_pct > 0.10 ~ v13_col_sn_2018$v2elsnlfc_15, TRUE ~ NA_real_)) %>%
  #"more free and fair subnational elections"
  mutate(elsnmr_15 = case_when(
    MOV_pct > 0.10 ~ v13_col_sn_2018$v2elsnmrfc_15, TRUE ~ NA_real_)) %>%
  #"stronger civil liberties" 
  mutate(clrgst_15 = case_when(
    MOV_pct > 0.10 ~ v13_col_sn_2018$v2clrgstch_15, TRUE ~ NA_real_)) %>%
  #"weaker civil liberties" 
  mutate(clrgwk_15 = case_when(
    MOV_pct > 0.10 ~ v13_col_sn_2018$v2clrgwkch_15, TRUE ~ NA_real_))

#Q16: in "areas where support [10% MOV] for the national ruling party is weak" there are ...
RP_15 <- RP_15 %>% 
  #"less free and fair subnational elections" 
  mutate(elsnl_16 = case_when(
    MOV_pct < -0.10 ~ v13_col_sn_2018$v2elsnlfc_16, TRUE ~ NA_real_)) %>%
  #"more free and fair subnational elections"
  mutate(elsnmr_16 = case_when(
    MOV_pct < -0.10 ~ v13_col_sn_2018$v2elsnmrfc_16, TRUE ~ NA_real_)) %>%
  #"stronger civil liberties" 
  mutate(clrgst_16 = case_when(
    MOV_pct < -0.10 ~ v13_col_sn_2018$v2clrgstch_16, TRUE ~ NA_real_)) %>%
  #"weaker civil liberties" 
  mutate(clrgwk_16 = case_when(
    MOV_pct < -0.10 ~ v13_col_sn_2018$v2clrgwkch_16, TRUE ~ NA_real_))

RP_15 <- RP_15[complete.cases(RP_15$FID), ] # remove NA case



```

```{r, V18_15-16_export, include=FALSE, echo=FALSE}

# Export the final merge to a .csv file that can be read into ArcMaps
RP18_15t16 <- RP_15 %>% 
  select(1:3|7|11:19) %>%
  rename(FID_15t16 = FID) %>%
  rename(Petro_15t16 = Petro) %>%
  rename(Duque_15t16 = Duque) %>%
  rename(vtotal_15t16 = vtotal) %>%
  rename(MOV18_15t16 = MOV_pct) %>%
  mutate(MOV18_15t16 = round(MOV18_15, 5))
#Belen de Bajira (Choco) does not have an FID number, which causes problems in ArcMap when merging. This municipality was only recently inaugurated.
RP18_15t16 <- RP18_15t16[complete.cases(RP18_15t16$FID_15t16), ]
# Export data to a csv file that will be joined to map
write.csv(RP18_15t16, file = "G:/My Drive/git/snvdem24/data/geospatial/2018pmq/15-16_RulingParty/RP18_15-16.csv", row.names = FALSE)

```

### Write shapefile for sf

```{r, sf_write, include=FALSE, echo=FALSE}
st_write(col, "G:/My Drive/git/snvdem24/data/geospatial/2018pmq/MGN_2018r.shp")

```

## Future research: Spatial analysis of democracy and development outcomes

(See more in Patrick's Causal Inference code)

### Education

-   coverage (feasible enrollment rate, sessions)

-   infrastructure (establishments)

-   test scores (Saber11 in math and reading)

-   others (e.g., student-teacher ratios)

### Health

-   Mortality (infant mortality, 1000 inhabitants)

-   Disease prevalence (Dengue, tuberculosis, HIV)

-   Vaccination (pentavalent, triple viral)

### Covariates

Interested in both time-variant and time-invariant information.

-   Time-variant: Population and density (rural and total), PIB per
    capita, Gini

-   Time-invariant: elevation (proxy for difficult access), distance
    from state capital

# Appendix

## Subnational V-Dem responses (nominal)

-   0: Rural. (0=No, 1=Yes) [v2elsnlfc_0]
-   1: Urban. (0=No, 1=Yes) [v2elsnlfc_1]
-   2: Areas that are less economically developed. (0=No, 1=Yes)
    [v2elsnlfc_2]
-   3: Areas that are more economically developed. (0=No, 1=Yes)
    [v2elsnlfc_2]
-   4: Inside the capital city. (0=No, 1=Yes) [v2elsnlfc_4]
-   5: Outside the capital city. (0=No, 1=Yes) [v2elsnlfc_5]
-   6: North. (0=No, 1=Yes) [v2elsnlfc_6]
-   7: South. (0=No, 1=Yes) [v2elsnlfc_7]
-   8: West. (0=No, 1=Yes) [v2elsnlfc_8]
-   9: East. (0=No, 1=Yes) [v2elsnlfc_9]
-   10: Areas of civil unrest (including areas where insurgent groups
    are active). (0=No, 1=Yes) [v2elsnlfc_10]
-   11: Areas where illicit activity is widespread. (0=No, 1=Yes)
    [v2elsnlfc_11]
-   12: Areas that are very sparsely populated. (0=No, 1=Yes)
    [v2elsnlfc_12]
-   13: Areas that are remote (difficult to reach by available
    transportation, for example). (0=No, 1=Yes) [v2elsnlfc_13]
-   14: Areas where there are indigenous populations. (0=No, 1=Yes)
    [v2elsnlfc_14]
-   15: Areas where the national ruling party or group is strong. (0=No,
    1=Yes) [v2elsnlfc_15]
-   16: Areas where the national ruling party or group is weak. (0=No,
    1=Yes) [v2elsnlfc_16]
-   17: Areas that were subject to a longer period of foreign rule.
    (0=No, 1=Yes) [v2elsnlfc_17]
-   18: Areas that were subject to a shorter period of foreign rule.
    (0=No, 1=Yes) [v2elsnlfc_18]
-   19: Areas that were recently subject to foreign rule. (0=No, 1=Yes)
    [v2elsnlfc_19]
-   20: Areas that have not recently been subject to foreign rule.
    (0=No, 1=Yes) [v2elsnlfc_20]
-   21: None of the above. (0=No, 1=Yes) [v2elsnlfc_21]

Shapefile
